biological data preperation

### 2019-01-16 src/biological_data_prep.Rmd
glorys bottom info added to presence data (but bottom layer no is not good)

>>>>

# glorys bathymetric data

so i just found that glorys also has a bathymetric data layer...

First you need to put the observations (data_aea) into a spatialpointsdataframe.
```{r}
xy <- data_aea[ ,c(19,20)] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
data_aea_sp <- SpatialPointsDataFrame(coords = xy, data = data_aea, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
```

then covert the netcdf with the bathy layer (GLO-MFC_001_025_bathy.nclonlat.nc) to a brick and project to the albers equal area
```{r}
bathyrast <- raster("../data/env/GLO-MFC_001_025_bathy.nclonlat.nc") #because it is a single level file
aea <- raster("../output/env/aea.tif") 
bathyrast <- projectRaster(bathyrast, aea)
plot(bathyrast)
```

now extract and add to data_aea
```{r}
data_aea_sp$bottom_depth_glorys <- extract(x = bathyrast, y = data_aea_sp)
head(data_aea_sp)
```

ok great - now let's do the same thing for the bottom_layer_no (yes Glorys has a layer for this as well!)

 covert the netcdf with the bottom layer nor (GLO-GLO-MFC_001_025_bottomlayerno.nclonlat.nc) to a raster and project to the albers equal area
```{r}
botlayernorast <- raster("../data/env/GLO-MFC_001_025_bottomlayerno.nclonlat.nc") #because it is a single level file
botlayernorast <- projectRaster(botlayernorast, aea)
plot(botlayernorast)
```

now extract and add to data_aea_sp, but this time also change the sp to a normal dataframe and save as a .csv
```{r}
data_aea_sp$bottom_layerno_glorys <- extract(x = botlayernorast, y = data_aea_sp)
write.csv(data_aea_sp, file = "presence_points_without_envdata_relooped_glbathy.csv", row.names = FALSE) # change the path to where ever you would like to save it
data_aea <- as.data.frame(data_aea_sp) # this coverts the 'spatialpoints dataframe' to a 'normal dataframe'
head(data_aea)
```

ok bottom layer no is not right
